---
title: "R Final Project"
author: "Ei Phyu Sin Win | Htet Khant Linn | Khant Razar Kyaw | Khine Nwe Lin"
date: "2025-12-02"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Installation

```{r}
library(tidyverse)
library(tidyr)
library(stringr)
library(lubridate)


```

# Data Loading

## Loading each monthly sales

```{r}
jan_sales <- read_csv("data/raw/2019-sales-monthly/201901-sales.csv")
head(jan_sales)
```

```{r}
feb_sales <- read_csv("data/raw/2019-sales-monthly/201902-sales.csv")
head(feb_sales)
```

```{r}
mar_sales <- read_csv("data/raw/2019-sales-monthly/201903-sales.csv")
head(mar_sales)
```

```{r}
apr_sales <- read_csv("data/raw/2019-sales-monthly/201904-sales.csv")
head(apr_sales)
```

```{r}
may_sales <- read_csv("data/raw/2019-sales-monthly/201905-sales.csv")
head(may_sales)
```

```{r}
jun_sales <- read_csv("data/raw/2019-sales-monthly/201906-sales.csv")
head(jun_sales)
```

```{r}
jul_sales <- read_csv("data/raw/2019-sales-monthly/201907-sales.csv")
head(jul_sales)
```

```{r}
aug_sales <- read_csv("data/raw/2019-sales-monthly/201908-sales.csv")
head(aug_sales)
```

```{r}
sep_sales <- read_csv("data/raw/2019-sales-monthly/201909-sales.csv")
head(sep_sales)
```

```{r}
oct_sales <- read_csv("data/raw/2019-sales-monthly/201910-sales.csv")
head(oct_sales)
```

```{r}
nov_sales <- read_csv("data/raw/2019-sales-monthly/201911-sales.csv")
head(nov_sales)
```

```{r}
dec_sales <- read_csv("data/raw/2019-sales-monthly/201912-sales.csv")
head(dec_sales)
```

## Combine the 12 months data using bind_rows

```{r}
all_sales_2019_combine <- bind_rows(jan_sales, feb_sales, mar_sales,
                            apr_sales, may_sales, jun_sales,
                            jul_sales, aug_sales, sep_sales,
                            oct_sales, nov_sales, dec_sales)

head(all_sales_2019_combine)
```


# EDA Process

## Clean the Data

### Check the Table Info

```{r}
glimpse(all_sales_2019_combine)
```

### Counts the Number of Missing Cells (NA)

```{r}
total_NA_cells <- sum(is.na(all_sales_2019_combine))
cat("Total no. of missing cells: ", total_NA_cells, "\n")

```

### Counts the Number of Rows with at Least One NA

```{r}
total_NA_rows <- sum(!complete.cases(all_sales_2019_combine))
cat("Total No. of missing rows: ", total_NA_rows, "\n")
```

### Counting Total No. of Rows Before Dropping

```{r}
row_before <- nrow(all_sales_2019_combine)
cat("Number of rows before dropping NA: ", row_before, "\n")
```

### Dropping NA Rows

```{r}
all_sales_2019 <- drop_na(all_sales_2019_combine)
# counting total no. of rows after dropping
row_after <- nrow(all_sales_2019)
cat("Number of rows after dropping NA: ", row_after, "\n")

# if this match with the total no. of missing row, we can validate it
cat("No. of dropped rows: ", row_before - row_after)
```

### Removing Duplicated Header Rows

```{r}
all_sales_2019 <- all_sales_2019 %>%
  # this remove the no. of repeated header rows
  filter(`Order ID` != "Order ID") %>%
  mutate(
    OrderID = `Order ID`,
    QuantityOrdered = as.integer(`Quantity Ordered`),
    PriceEach = as.numeric(`Price Each`),
    OrderDate = `Order Date`,
    PurchaseAddress = `Purchase Address`,
    ZIP = str_extract(PurchaseAddress, "\\d{5}$")
  ) %>%
  select(OrderID, Product, QuantityOrdered, PriceEach, OrderDate, PurchaseAddress, ZIP)

row_after_cleaning_repeated_header <- nrow(all_sales_2019)

cat("Total no. of repeated header rows that was dropped: ", row_after - row_after_cleaning_repeated_header)
```

### Check the Table Info & Exporting

```{r}
glimpse(all_sales_2019)

write_csv(all_sales_2019, "data/processed/all_sales_2019.csv")
```

---

## Data Manipulation
### Separating PurchaseAddress into 3 Columns (Street, City and State_ZIP)

```{r}
all_sales_2019_splitted_and_formatted <- separate(all_sales_2019, PurchaseAddress, into = c("Street", "City", "State_ZIP"), sep = ", ")
```


### Add TotalSales & Separate City and ZIP columns
```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>%
  mutate(
    TotalSales = QuantityOrdered * PriceEach,
    City = str_c(City, " (", str_extract(State_ZIP, "^[A-Z]{2}"), ")"),
    ZIP = str_extract(State_ZIP, "\\d{5}$")
    )
```

### Adding OrderDataTime, Month, and Hour Columns
```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>%
  mutate(
    OrderDateTime = mdy_hm(OrderDate),
    Month = month(OrderDateTime, label = TRUE),
    Hour = hour(OrderDateTime)
  )
```

### Select and Check the Clean Data
```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>% 
  select(
    OrderID, Product, QuantityOrdered, PriceEach, TotalSales, Street, City, ZIP, Month, Hour, OrderDateTime
    )

glimpse(all_sales_2019_splitted_and_formatted)
```

### Exporting Clean Data into CSV
```{r}
# write_csv(all_sales_2019_splitted_and_formatted, "data/processed/all_sales_2019_v2.csv")

```
