---
title: "2025-11-09-R-Final-Project-V3"
author: "Ei Phyu Sin Win, Htet Khant Linn, Khant Razar Kyaw, Khine Nwe Lin"
date: "2025-12-03"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Installation

```{r}
library(tidyverse)
library(tidyr)
library(stringr)
library(lubridate)
library(readr)
library(ggplot2)
library(dplyr)
library(scales)
library(forcats)
library(forecast)
library(tseries)

```


# Data Loading

## Loading each monthly sales

```{r}
jan_sales <- read_csv("data/raw/2019-sales-monthly/201901-sales.csv")
head(jan_sales)
```

```{r}
feb_sales <- read_csv("data/raw/2019-sales-monthly/201902-sales.csv")
head(feb_sales)
```

```{r}
mar_sales <- read_csv("data/raw/2019-sales-monthly/201903-sales.csv")
head(mar_sales)
```

```{r}
apr_sales <- read_csv("data/raw/2019-sales-monthly/201904-sales.csv")
head(apr_sales)
```

```{r}
may_sales <- read_csv("data/raw/2019-sales-monthly/201905-sales.csv")
head(may_sales)
```

```{r}
jun_sales <- read_csv("data/raw/2019-sales-monthly/201906-sales.csv")
head(jun_sales)
```

```{r}
jul_sales <- read_csv("data/raw/2019-sales-monthly/201907-sales.csv")
head(jul_sales)
```

```{r}
aug_sales <- read_csv("data/raw/2019-sales-monthly/201908-sales.csv")
head(aug_sales)
```

```{r}
sep_sales <- read_csv("data/raw/2019-sales-monthly/201909-sales.csv")
head(sep_sales)
```

```{r}
oct_sales <- read_csv("data/raw/2019-sales-monthly/201910-sales.csv")
head(oct_sales)
```

```{r}
nov_sales <- read_csv("data/raw/2019-sales-monthly/201911-sales.csv")
head(nov_sales)
```

```{r}
dec_sales <- read_csv("data/raw/2019-sales-monthly/201912-sales.csv")
head(dec_sales)
```

## Combine the 12 months data using bind_rows

```{r}
all_sales_2019_combine <- bind_rows(jan_sales, feb_sales, mar_sales,
                            apr_sales, may_sales, jun_sales,
                            jul_sales, aug_sales, sep_sales,
                            oct_sales, nov_sales, dec_sales)

head(all_sales_2019_combine)
```


# EDA Process

## Clean the Data

### Check the Table Info

```{r}
glimpse(all_sales_2019_combine)
```

### Counts the Number of Missing Cells (NA)

```{r}
total_NA_cells <- sum(is.na(all_sales_2019_combine))
cat("Total no. of missing cells: ", total_NA_cells, "\n")

```

### Counts the Number of Rows with at Least One NA

```{r}
total_NA_rows <- sum(!complete.cases(all_sales_2019_combine))
cat("Total No. of missing rows: ", total_NA_rows, "\n")
```

### Counting Total No. of Rows Before Dropping

```{r}
row_before <- nrow(all_sales_2019_combine)
cat("Number of rows before dropping NA: ", row_before, "\n")
```

### Dropping NA Rows

```{r}
all_sales_2019 <- drop_na(all_sales_2019_combine)
# counting total no. of rows after dropping
row_after <- nrow(all_sales_2019)
cat("Number of rows after dropping NA: ", row_after, "\n")

# if this match with the total no. of missing row, we can validate it
cat("No. of dropped rows: ", row_before - row_after)
```

### Removing Duplicated Header Rows

```{r}
all_sales_2019 <- all_sales_2019 %>%
  # this remove the no. of repeated header rows
  filter(`Order ID` != "Order ID") %>%
  mutate(
    OrderID = `Order ID`,
    QuantityOrdered = as.integer(`Quantity Ordered`),
    PriceEach = as.numeric(`Price Each`),
    OrderDate = `Order Date`,
    PurchaseAddress = `Purchase Address`,
    ZIP = str_extract(PurchaseAddress, "\\d{5}$")
  ) %>%
  select(OrderID, Product, QuantityOrdered, PriceEach, OrderDate, PurchaseAddress, ZIP)

row_after_cleaning_repeated_header <- nrow(all_sales_2019)

cat("Total no. of repeated header rows that was dropped: ", row_after - row_after_cleaning_repeated_header)
```

### Check the Table Info & Exporting

```{r}
glimpse(all_sales_2019)

write_csv(all_sales_2019, "data/processed/all_sales_2019.csv")
```

## Data Manipulation

### Separating PurchaseAddress into 3 Columns (Street, City and State_ZIP)

```{r}
all_sales_2019_splitted_and_formatted <- separate(all_sales_2019, PurchaseAddress, into = c("Street", "City", "State_ZIP"), sep = ", ")
```

### Add TotalSales & Separate City and ZIP columns

```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>%
  mutate(
    TotalSales = QuantityOrdered * PriceEach,
    City = str_c(City, " (", str_extract(State_ZIP, "^[A-Z]{2}"), ")"),
    ZIP = str_extract(State_ZIP, "\\d{5}$")
    )
```

### Adding OrderDataTime, Month, and Hour Columns

```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>%
  mutate(
    OrderDateTime = mdy_hm(OrderDate),
    Month = month(OrderDateTime, label = TRUE),
    Hour = hour(OrderDateTime)
  )
```

### Select and Check the Clean Data

```{r}
all_sales_2019_splitted_and_formatted <- all_sales_2019_splitted_and_formatted %>% 
  select(
    OrderID, Product, QuantityOrdered, PriceEach, TotalSales, Street, City, ZIP, Month, Hour, OrderDateTime
    )

glimpse(all_sales_2019_splitted_and_formatted)
```

### Exporting Clean Data into CSV

```{r}
# write_csv(all_sales_2019_splitted_and_formatted, "data/processed/all_sales_2019_v2.csv")

```

# EDA Process

## Importing Data

```{r}
all_sales_2019 <- read_csv("data/processed/all_sales_2019_v2.csv")

# Check data structure and Null Value.
head(all_sales_2019)
glimpse(all_sales_2019)

```

## Calcuating Summary Statistics

```{r}
cat("=== General Summary of Data ===\n")
summary(all_sales_2019 %>% select(QuantityOrdered, PriceEach, TotalSales))
```

## Checking Null Value

```{r}
sum(is.na(all_sales_2019))
```

# Question 1: What was the best month for sales? How much was earned that month?

```{r}
total_sales_each_month <- all_sales_2019 %>%
  group_by(Month) %>%
  summarize(TotalSales = sum(TotalSales)) %>%
  arrange(desc(TotalSales))

# write_csv(total_sales_each_month, "outputs/tables/total_sales_each_month.csv")

month_order = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

# ordering month for visual
total_sales_each_month <- total_sales_each_month %>%
  mutate(Month = factor(Month, levels = month_order))

vis_1 <- ggplot(total_sales_each_month, aes(x = Month, y = TotalSales)) +
  geom_col() + 
  labs(title = "Total Sales by Month",
       x = "Month",
       y = "Total Sales in USD ($)")

vis_1 <- vis_1 + 
  geom_text(aes(label = scales::dollar(TotalSales, scale = 1/1e6, suffix = "M")),
            vjust = -0.5,
            size = 2.5) +
  scale_y_continuous(
    labels = scales::dollar_format(scale = 1/1e6, suffix = "M"),
    expand = expansion(add = c(0, 1e6))
    ) +
  theme_minimal()

# ggsave("outputs/figures/total_sales_by_month.jpg", plot = vis_1)
print(vis_1)



best_month_for_sales <- total_sales_each_month %>%
  slice_max(TotalSales, n = 1)

cat("The best month for sales is", best_month_for_sales$Month, 
    "with the total sales of", 
    paste0("$", best_month_for_sales$TotalSales, "."),
    "\n")
```


# Question 2: What city has the highest sales?

```{r}
total_sales_each_city <- all_sales_2019 %>%
  group_by(City) %>%
  summarize(TotalSales = sum(TotalSales)) %>%
  arrange(desc(TotalSales))

# write_csv(total_sales_each_city, "outputs/tables/total_sales_each_city.csv")

vis_2 <- ggplot(total_sales_each_city, aes(TotalSales, City)) +
  geom_col()

vis_2 <- vis_2 + 
  labs(
    title = "Total Sales per City",
    x = "Total Sales n USD ($)",
    y = "City Name"
  )+
  geom_text(aes(label = scales::dollar(TotalSales, scale = 1/1e6, suffix = "M")),
                   hjust = -0.2,
                   size = 2.5) +
  scale_x_continuous(
    label = scales::dollar_format(scale = 1/1e6, suffix = "M"),
     expand = expansion(add = c(0, 1e6))
    ) + 
  theme_minimal()

# ggsave("outputs/figures/total_sales_per_city.jpg", plot = vis_2)
print(vis_2)

city_with_highest_sales <- total_sales_each_city %>%
  slice_max(TotalSales, n = 1)


cat("The city with highest sales is", city_with_highest_sales$City, 
    "with the total sales of", 
    paste0("$", city_with_highest_sales$TotalSales, "."),
    "\n")
```


# Question 3: What time should we display advertisements to maximize likehood of customers buying products?

```{r}
glimpse(all_sales_2019)

sales_by_hours <- all_sales_2019 %>%
  group_by(Hour) %>%
  summarize(TotalQuantitySales = sum(QuantityOrdered))

# write_csv(sales_by_hours, "outputs/tables/sales_by_hours.csv")

vis_3 <- ggplot(sales_by_hours, aes(x = Hour, y = TotalQuantitySales)) +
  geom_line(size = 1) +
  labs(
    title = "Total Sales by Hours",
    y = "Total Sales in USD ($)",
    x = "Hour of the Day (24 Hrs)"
  ) +
  scale_x_continuous(breaks = seq(0, 24, by = 4)) +
  scale_y_continuous(breaks = seq(0, 17500, by = 2500)) +
  theme_minimal()

# ggsave("outputs/figures/total_sales_by_hours.jpg", plot = vis_3)
print(vis_3)
```


# Question 4: What products are most often sold together?

```{r}
duplicated_OrderID <- all_sales_2019 %>%
  group_by(OrderID) %>%
  filter(n() > 1) %>% 
  select(OrderID, Product) %>%
  ungroup()

products_sold_together <- duplicated_OrderID %>%
  group_by(OrderID) %>%
#  count(sort = TRUE) %>%
  summarize(ProductPairs = paste(sort(Product), collapse = ", ")) %>%
  ungroup()

products_sold_together_count <- products_sold_together %>%
  count(ProductPairs, sort = TRUE) %>%
  rename(TotalSales = n)

top_10_products_sold_together <- head(products_sold_together_count, n = 10)

print(top_10_products_sold_together)
# write_csv(products_sold_together_count, "outputs/tables/products_sold_together.csv")
```


# Question 5: What product sold the most? Why?

```{r}
single_product_sold <- all_sales_2019 %>%
  group_by(Product) %>%
  count(wt = QuantityOrdered, sort = TRUE) %>%
  rename(TotalUnitSold = n)

most_single_product_sold <- single_product_sold %>%
  head(1)


cat(paste0("The most sold product is ", most_single_product_sold$Product,
           " with the total sales of ", most_single_product_sold$TotalUnitSold,
           " items. \n"))

# write_csv(single_product_sold, "outputs/tables/product_sales.csv")

vis_4 <- ggplot(single_product_sold, aes(TotalUnitSold, fct_reorder(Product, TotalUnitSold))) +
  geom_col() +
  labs(
    title = "Product Sales",
    x = "Number of Units Sold",
    y = "Product Name"
  ) +
  theme_minimal()

# ggsave("outputs/figures/product_sales.jpg", plot = vis_4)
print(vis_4)
```

```{r}
# Get Quantity and Price by Product
trend <- all_sales_2019 %>%
  group_by(Product) %>%
  summarize(Qty = sum(QuantityOrdered), Price = mean(PriceEach))

# Calculate scaling factor
coeff <- max(trend$Qty) / max(trend$Price)

# 
ggplot(trend, aes(x = Product)) +
  geom_col(aes(y = Qty)) + 
  geom_line(aes(y = Price * coeff, group = 1), color = "red", size = 0.5) + 
  geom_point(aes(y = Price * coeff), color = "green", size = 1) + 
  scale_y_continuous(
    name = "Quantity", 
    sec.axis = sec_axis(~ . / coeff, name = "Mean Price ($)")
  ) +
  coord_flip() +  # rotates the chart
  theme_minimal()

```

From the above viausl we can see **inverse correlation**, which is cheap products (like AAA Batteries and Charging Cables) sell in very high quantities, while expensive products (like Laptops and Monitors) sell in lower quantities.
However, there are some exception such as Even though the Mac book Pro Laptop is the most expensive item (the green dot is furthest to the right), it still sells more units than cheaper home appliances like the LG Washing Machine.
This shows that customer demand for laptops is much higher than for appliances, regardless of the high price.

This proves that while low price usually drives high volume, strong brand popularity (like Apple) or high necessity (like work laptops) can still drive sales even at high prices.




# Building Time series model

```{r}
# all_sales_2019 <- read_csv("data/processed/all_sales_2019_v2.csv")
```

## aggregate sales by day

```{r}
daily_sales <- all_sales_2019 %>%
  mutate(Date = as.Date(OrderDateTime)) %>% # Date format
  group_by(Date) %>%
  summarize(TotalSales = sum(TotalSales)) %>%
  arrange(Date) # data is ordered chronologically

```

## create a Time Series Object

```{r}
# use frequency = 7 >>> weekly seasonality
sales_ts2 <- ts(daily_sales$TotalSales, frequency = 7)
```

# analyze Differencing

```{r}
sales_diff <- diff(sales_ts2)
# we look at the Differenced data (Day 2 - Day 1)
```

# plot ACF and PACF

```{r}
ggtsdisplay(sales_diff, main="ACF & PACF of Differenced Sales")
```

The ACF plot shows clear spikes at Lag 7, 14, and 21. This confirm a strong weekly cycle, meaning sales consistently peak on specific days (likely weekends).
The top chart shows the data fluctuating around zero without drifting up or down, which means the data is stationary after differencing.

# MODEL BUILDING

## Model A: (ARIMA 0,1,1)

```{r}
model_a <- Arima(sales_ts2, order = c(0,1,1))
```

## Model B: (ARIMA 1,1,1)

```{r}
model_b <- Arima(sales_ts2, order = c(1,1,1))
# checks if yesterday's sales value directly predicts today's
```

## Model C: (SARIMA 1,1,1)(1,0,0)[7]

```{r}
model_c <- Arima(sales_ts2, 
                 order = c(1,1,1), 
                 seasonal = list(order = c(1,1,1), period = 7))
# looks for a 7-day (weekly) pattern
```

# Compare Models

```{r}
cat("AIC Comparison (note: Lower is Better):\n")
cat("Model A (0,1,1):", AIC(model_a), "\n")
cat("Model B (1,1,1):", AIC(model_b), "\n")
cat("Model C (Seasonal):", AIC(model_c), "\n")
```
We tested three models and Model C (Seasonal ARIMA) was the most accurate because it explicitly account for the weekly sales pattern. It achieve the lowest AIC score (7833.5), which is significantly better than the non-seasonal models (7950.2 and 7952.0), indicating much tighter fit to the real data.

# Visual Comparison of Forecasts

```{r}
autoplot(sales_ts2) +
  autolayer(forecast(model_a, h=30), series="Model A (0,1,1)", PI=FALSE) +
  autolayer(forecast(model_b, h=30), series="Model B (1,1,1)", PI=FALSE) +
  autolayer(forecast(model_c, h=30), series="Model C (Seasonal)", PI=FALSE) +
  labs(title = "Comparing Straight Line vs. Seasonal Forecasts",
       y = "Daily Sales ($)") +
  theme_minimal()
```

# visualization with the uncertainty

```{r}
autoplot(sales_ts2) +
  autolayer(forecast(model_c, h=30), series="Model C (Seasonal)", PI=TRUE) +
  labs(title = "Comparing Straight Line vs. Seasonal Forecasts",
       y = "Daily Sales ($)") +
  theme_minimal()
```


Unlike simple straight line, our final forecast predict that sales will continue to fluctuate week-by-week, preserving the historical "up and down" trend of the business. This is more realistic than the simple linear regression we previously tested, which only explained about 27% of the data variation.


# Building Linear Model (Archive Part - only for review purpose) 
**Please note that we only add this part to show what we have done before Professor feedback during the presentation. We have changed our model to Time Series as mentioned above according to professor feedback**

## 1. Prepare Data: Aggregate sales by day

```{r}
daily_sales2 <- all_sales_2019 %>%
  mutate(Date = as.Date(OrderDateTime, format = "%m/%d/%y %H:%M")) %>%
  group_by(Date) %>%
  summarize(TotalSales = sum(QuantityOrdered * PriceEach)) %>%
  mutate(DayIndex = as.numeric(Date - min(Date) + 1)) # Create a numeric time variable
```

## 2. Build Linear Model (Simple Linear Regression)

Independent Variable (X): Day of the year (Time).
Dependent Variable (Y): Total Daily Sales.
Goal: Determine if sales are significantly increasing or decreasing as the year progresses.

```{r}
# Model: TotalSales = Intercept + Slope * DayIndex
sales_model <- lm(TotalSales ~ DayIndex, data = daily_sales2)
```

## 3. Model Evaluation

```{r}
model_summary <- summary(sales_model)
print(model_summary)
```

**Model Evaluation** **Intercept(70834.64)**: The model estimates that on Day 0 (the very start of the year), the business would have made about \$70834.64 in sales.
**DayIndex(127.55)**: For every single day that passes in 2019, the total sales increased by an average of \$127.55.Since this number is positive, the business is growing daily throughout the year.

**P-Value**: The P-Value is extremely small, showing statistically significant.
Because the value is less than 0.05, we are confident that the increase in sales is real and not just due to random luck.

**R-Squared Value**

```{r}
# Extract R-squared to see how well the date explains sales variance
r_squared <- model_summary$r.squared
cat("R-squared:", r_squared, "\n")
```

While there is a clear upward trend (sales increasing by \~\$127/day), the date alone only explains about 27% of the sales volume.
The remaining 73% of variation is *likely due to weekly cycles (weekends vs weekdays) and holiday spikes*.

## 4. Visualization with Regression Line

```{r}
ggplot(daily_sales2, aes(x = Date, y = TotalSales)) +
  geom_point(alpha = 0.5, color = "navyblue") +      # Scatter plot of actual data
  geom_smooth(method = "lm", color = "red") +    # The regression line
  labs(title = "Daily Sales Trend (2019)",
       subtitle = paste("R-squared:", round(r_squared, 4)),
       x = "Date",
       y = "Total Sales ($)") +
  theme_minimal()

# ggsave("outputs/figures/sales_trend_regression.png")
```


